---
title: "House Price Predictions - Predicitive Model (Part2)"
author: "Christine Utendorf"
date: "23 May 2019"
output: html_document
---

## 0. Sources and data loading

#### Notebook loading
Besides several libraries used for data exploration and visualization, the library loding also includes MASS, caret and ranger for the machine learning models.

All self-programed functions such as transforming or splitting a dataset are saved within the R notbook called functions.

Furthermore, several metrics that are not default for machine learning R packages are created as functions within the R notebook metrics. 
```{r, warning=FALSE, message=FALSE}
source('notebooks/libraries.R')
source('notebooks/functions.R')
source('notebooks/metrics.R')
```

The work is divided into two R markdowns. The first one only focusses on data exploration and preparation and the second one on modeling. This is due to the fact that each markdown takes some while to be created and with this division the whole data exploration does not need to be computed again while modeling.

#### Data loading
From the data preparation and exploration part the prepared train and test data sets are loaded.

```{r}
train_data<-fread('data/data_train_dummy.csv', stringsAsFactors = F)
test_data<-fread('data/data_test_dummy.csv', stringsAsFactors = F)
```

#### Splitting dataset
Since our test set is blind and we want to compare models when tuning them, a list object is created that not only stores the test data set but also a train data set (80% of train data) and a holdout set (20% of train data). Furthermore our target variable price is set and all integer variables are turned into numerical ones.

We store the id and the price data for the holdout in a seperate data set in order to calculate later on the metrics. Furthermore also the ids of the blind test set are stored in a seperate dataset to later on store the final predictions for submission.

Date and id and then excluded from all datasets since there were severe problems when trying to predict with the models.

```{r}
whole_data = f_split(train_data, test_data, testsize = 0.2, seed = 1)
whole_data<-lapply(whole_data, function(x){
  return(x[, which(sapply(x, is.integer)):=lapply(.SD, as.numeric), .SDcols=sapply(x,is.integer)])
})

#df_test<-whole_data$holdout[,c("id", "price")]
df_test<-whole_data$holdout[, .(id=1:.N, price)]
df_pred<-whole_data$test[,c("id")]

whole_data$train$id = NULL
whole_data$test$id = NULL
whole_data$holdout$id = NULL

formula<-as.formula(price~.)
```


## 1. Linear model
To start with a linear model is used to predict the continous variable.

```{r, warning=FALSE, message=FALSE}
lm <- lm(formula = formula, 
                 data=whole_data$train)
summary(lm)

test_lm<-predict(lm, newdata = whole_data$holdout)

df_test<-cbind(df_test, test_lm)

str(df_test)

ggplot(melt(df_test, id.vars = 'id'), aes(x=id,y=value, colour=variable))+
  geom_point(alpha=0.65)+ylim(0,4000000)+xlab('')+ylab('$')+
  ggtitle('Linear Regression - Test prediction')+
  scale_colour_manual(values = c('black','red'))


rmse_lm<-rmse(real=whole_data$holdout$price, predicted = test_lm)
mae_lm<-mae(real=whole_data$holdout$price, predicted = test_lm)
mape_lm<-mape(real=whole_data$holdout$price, predicted = test_lm)
mape_lm
```
